---
title: "tidy-model"
author: "Kevin Guan"
date: "2025-02-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set
library('tidyverse') ; library('modeldata')
```
## Introduction

This R Markdown document demonstrates my abilities to use models for data analysis using a data set collected on crickets.

```{r load data, include = FALSE}
data(crickets, package = "modeldata")
names(crickets)
```

## What is that sound?

In this report, we examine what factors predict a cricket's chirp rate.
The data set contains a total of 31 observations from 2 species of crickets. The measured temperature ranged from 17.2 to 30.4 degrees Celsius, and the mean rate of chirping was 72.89 chirps per min. 

```{r summary, echo= FALSE}
summary(crickets)
ggplot(crickets, aes(x = rate)) +
  geom_histogram(bins = 8) +
  ggtitle("Distribution of the chirp rates of crickets") +
  xlab('Chirp rate (per min.)')
```

## Temperature affects chirp rate
```{r temp, echo=FALSE}
ggplot(crickets, aes(x = temp, y = rate)) +
  geom_point() +
  geom_smooth(method = 'lm') +
  ggtitle("Plot of temperature and chirp rate") +
  ylab('Chirp rate (per min.)') +
  xlab('Temperature (Celsius)')
```

Based on a scatter plot of temperature and chirp rate, we can observer that as temperature increases, so does the chirp rate.
